---
title: My Example Computed Manuscript
subtitle: Created in Rmarkdown
titlerunning: Example computed manuscript

authors: 
- name: Jeffrey M. Perkel
  address: Springer Nature, 1 New York Plaza, New York, NY
  email: jeffrey.perkel@nature.com
  
abstract: |
  A mock computed manuscript created in RStudio using {Rmarkdown}. The {Bookdown} and {Rticles} packages were used to output the text in Springer Nature's desired manuscript format. 

bibliography: bibliography.bib
biblio-style: spbasic
# bibstyle options spbasic(default), spphys, spmpsci
output: 
  bookdown::pdf_book:
    base_format: rticles::springer_article
    extra_dependencies: ["booktabs"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      echo = FALSE)
```

```{r load-libraries, include=FALSE}
# load libraries
library(tidyverse)
library(ggbeeswarm)
library(bookdown)
```

# Introduction {#intro}

"Literate programming" is a style of programming that uses computational notebooks to weave together code, explanatory text, data and results into a single document, enhancing scientific communication and computational reproducibility.[@shen2014; @perkel2018a; @perkel2018]

Several platforms for creating such documents exist, including Jupyter, R Markdown and Observable notebooks.[@perkel2021] Typically, these documents interleave code and text 'blocks' to build a computational narrative. But some, including R Markdown and the JupyterBook extension to the Jupyter ecosystem, also allow authors to include and execute code "inline" -- that is, within the text itself.

This makes it possible to create fully executable manuscripts in which the document itself computes and inserts values and figures into the text rather than requiring authors to input them manually. This is in many ways the 'killer feature' of computed manuscripts: it circumvents the possibility that the author will enter an incorrect number, or forget to update a figure or value should new data arise.

In this manuscript, created in RStudio using the R Markdown language, we will create such an example.

# Results {#results}

## Inline computation {#sec:1}

Imagine we are analyzing data from a clinical trial. We have grouped subjects in three bins and measured the concentration of some metabolite. (These data are simulated.)

```{r initial-data}
# read in some initial data
df1 <- read_csv('data/example-data-1.csv')
```

Rather than analyzing those data and then copying the results into our manuscript, we can use the programming language `R` to do that in the manuscript itself. Simply enclose the code inside backticks, with the letter `r`. For instance, to calculate the circumference and area of a circle with radius *r* = 10, you could write "A = `` `r
pi * r^2` ``" and "C = `` `r
2 * pi * r` ``. Those evaluate to "A = **`r r = 10; round(pi * r^2, 3)`** and C = **`r r = 10; round(2 * pi * r, 3)`**".

Returning to our dataset, we have **`r nrow(df1)`** (simulated) subjects in our study (see Table \@ref(tab:show-table-1)). The average metabolite concentration is **`r round(mean(df1$conc), 2)`** (range: **`r min(df1$conc)`**-**`r max(df1$conc)`**). We have **`r df1 %>% filter(class == 'Group 1') %>% nrow()`** subjects in Group 1, **`r df1 %>% filter(class == 'Group 2') %>% nrow()`** subjects in Group 2, and **`r df1 %>% filter(class == 'Group 3') %>% nrow()`** in Group 3. (The numbers in **bold face type** throughout this document are computed values.)

```{r show-table-1}
tmp <- cbind(df1[1:33,], rep('|', 33), 
             df1[34:66,], rep('|', 33), 
             df1[67:99,])
names(tmp) <- c('ID', 'Class', 'Conc', '|', 'ID', 'Class', 'Conc', 
                '|', 'ID', 'Class', 'Conc')

knitr::kable(tmp, format = 'latex', booktabs = TRUE, 
             caption = "Initial subject data")
```

```{r new-data}
# read new dataset
df2 <- read_csv('data/example-data-2.csv')
```

## Incorporating new data {#sec:2}

Now suppose we get another tranche of data (Table \@ref(tab:show-table-2)). There are **`r nrow(df2)`** subjects in this new dataset. Their average concentration is **`r round(mean(df2$conc), 2)`** (range: **`r min(df2$conc)`**-**`r max(df2$conc)`**).

```{r show-table-2}
tmp <- cbind(df2[1:20,], rep('|', 20), 
             df2[21:40,], rep('|', 20), 
             df2[41:60,])
names(tmp) <- c('ID', 'Class', 'Conc', '|', 'ID', 'Class', 'Conc', 
                '|', 'ID', 'Class', 'Conc')

knitr::kable(tmp, format = 'latex', booktabs = TRUE, 
             caption = "New subject data")
```

```{r combine-tables}
# merge datasets
final_data <- rbind(df1, df2)
```

Combining the two datasets, we have a total of **`r nrow(final_data)`** subjects (Table \@ref(tab:show-table)). The revised average metabolite concentration is **`r round(mean(final_data$conc), 2)`** (range: **`r min(final_data$conc)`**-**`r max(final_data$conc)`**). We now have **`r final_data %>% filter(class == 'Group 1') %>% nrow()`** subjects in Group 1, **`r final_data %>% filter(class == 'Group 2') %>% nrow()`** subjects in Group 2, and **`r final_data %>% filter(class == 'Group 3') %>% nrow()`** in Group 3.

## Plotting the data {#sec:3}

We can also create and include figures during manuscript creation. Here we graph boxplots of our clinical trial data. The data are shown in Figure \@ref(fig:plot-data). Note that this figure number (as well as the table numbers    above) is automatically generated.

```{r plot-data, fig.cap="Metabolite concentration of clinical trial subjects", fig.height=3, fig.width=4}
# plot the data
final_data %>% 
  ggplot(aes(x = class, y = conc, color = class)) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(width = 0.25) + 
  xlab("") +
  ylab("Metabolite concentration") + 
  theme_minimal() +
  theme(legend.position = "none")

```

# Code {#code}

The following code was used to load, merge, and plot the (simulated) clinical trial data:

```{r show-code-1, echo=TRUE, eval=FALSE, ref.label='load-libraries'}
```

```{r show-code-2, echo=TRUE, eval=FALSE, ref.label='initial-data'}
```

```{r show-code-3, echo=TRUE, eval=FALSE, ref.label='new-data'}
```

```{r show-code-4, echo=TRUE, eval=FALSE, ref.label='combine-tables'}
```

```{r show-code-5, echo=TRUE, eval=FALSE, ref.label='plot-data'}
```

# References
