---
title: "My Example Computed Manuscript"
subtitle: Created in Rmarkdown
titlerunning: Example computed manuscript
date: "`r format(Sys.time(), '%d %b %Y %H:%M:%S %Z')`"
author: "Jeffrey M. Perkel, Technology Editor, Nature"
output:
  bookdown::html_document2: default
  pdf_document: default
  bookdown::word_document2: default
  bookdown::pdf_book:
    base_format: rticles::springer_article
    extra_dependencies: booktabs
abstract: "A mock computed manuscript created in RStudio using {Rmarkdown}. The {Bookdown}
  and {Rticles} packages were used to output the text in Springer Nature's desired
  manuscript format. \n"
bibliography: bibliography.bib
biblio-style: spbasic
authors:
- name: Jeffrey M. Perkel
  address: Springer Nature, 1 New York Plaza, New York, NY
  email: jeffrey.perkel@nature.com
csl: nature.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      echo = FALSE)
```

```{r load-libraries, include=FALSE}
# load libraries
library(tidyverse)
library(ggbeeswarm)
library(bookdown)
```

# Introduction {#intro}

"Literate programming" is a style of programming that uses computational notebooks to weave together code, explanatory text, data and results into a single document, enhancing scientific communication and computational reproducibility.[@shen2014; @perkel2018a; @perkel2018] (These references were added into the document using RStudio's integration with the open-source Zotero reference manager [@perkel2020] plus the [Better BibTeX](https://retorque.re/zotero-better-bibtex/) Zotero plugin.)

Several platforms for creating such documents exist.[@perkel2021] Typically, these documents interleave code and text 'blocks' to build a computational narrative. But some, including R Markdown, Observable, and the JupyterBook extension to the Jupyter ecosystem, also allow authors to include and execute code "inline" -- that is, within the text itself.

This makes it possible to create fully executable manuscripts in which the document itself computes and inserts values and figures into the text rather than requiring authors to input them manually. This is in many ways the 'killer feature' of computed manuscripts: it circumvents the possibility that the author will enter an incorrect number, or forget to update a figure or value should new data arise. Among other uses, that allows authors to automatically time-stamp their documents, or insert the current version number of the software they use into their methods. For instance, this document was built at **`r format(Sys.time(), "%d %b %Y %H:%M:%S %Z")`** and calls the following R packages: `{tidyverse}` ver. **`r packageVersion("tidyverse")`**, `{ggbeeswarm}` ver. **`r packageVersion("ggbeeswarm")`** and `{bookdown}` ver. **`r packageVersion("bookdown")`**.

In this manuscript, created in RStudio using the R Markdown language, we will demonstrate a more practical example. (A JavaScript version is available at [https://observablehq.com/d/7a08166446057936](https://observablehq.com/d/7a08166446057936).)

# Results {#results}

## Inline computation {#sec:1}

Imagine we are analyzing data from a clinical trial. We have grouped subjects in three bins and measured the concentration of some metabolite. (These data are simulated.)

```{r initial-data}
# read in some initial data
df1 <- read_csv('data/example-data-1.csv')
```

```{r radius}
# radius of a circle
r = 10
```

Rather than analyzing those data and then copying the results into our manuscript, we can use the programming language `R` to do that in the manuscript itself. Simply enclose the code inside backticks, with the letter `r`. For instance, to calculate the circumference and area of a circle with radius *r* = **`r r`**, you could write "A = `` `r
pi * r^2` `` and C = `` `r
2 * pi * r` ``". That evaluates to "A = **`r round(pi * r^2, 2)`** and C = **`r round(2 * pi * r, 2)`**".

Returning to our dataset, we have **`r nrow(df1)`** (simulated) subjects in our study (see Table \@ref(tab:show-table-1); see [`R/mock_data.R`](https://github.com/jperkel/computed_manuscript/blob/main/R/mock_data.R) in the GitHub repository for code to generate the mock dataset). Note the numbering of the tables, figures and sections in this document -- that's accomplished automatically thanks to the `{bookdown}` package. 
The average metabolite concentration in our dataset is **`r round(mean(df1$conc), 2)`** (range: **`r paste(min(df1$conc), max(df1$conc), sep = ' to ')`**). We have **`r df1 %>% filter(class == 'Group 1') %>% nrow()`** subjects in Group 1, **`r df1 %>% filter(class == 'Group 2') %>% nrow()`** subjects in Group 2, and **`r df1 %>% filter(class == 'Group 3') %>% nrow()`** in Group 3. (The numbers in **bold face type** throughout this document are computed values.)

```{r new-data}
# read new dataset
df2 <- read_csv('data/example-data-2.csv')
```

## Incorporating new data {#sec:2}

Now suppose we get another tranche of data (Table \@ref(tab:show-table-2)). There are **`r nrow(df2)`** subjects in this new dataset. Their average concentration is **`r round(mean(df2$conc), 2)`** (range: **`r paste(min(df2$conc), max(df2$conc), sep = ' to ')`**).

```{r combine-tables}
# merge datasets
final_data <- rbind(df1, df2)
```

Combining the two datasets, we have a total of **`r nrow(final_data)`** subjects. The revised average metabolite concentration is **`r round(mean(final_data$conc), 2)`** (range: **`r paste(min(final_data$conc), max(final_data$conc), sep = ' to ')`**). We now have **`r final_data %>% filter(class == 'Group 1') %>% nrow()`** subjects in Group 1, **`r final_data %>% filter(class == 'Group 2') %>% nrow()`** subjects in Group 2, and **`r final_data %>% filter(class == 'Group 3') %>% nrow()`** in Group 3. The concentration distribution for each group is shown graphically in Figure \@ref(fig:plot-data).

```{r plot-data, fig.cap="Metabolite concentration of clinical trial subjects", fig.height=3, fig.width=4}
# plot the data
p <- final_data %>% 
  ggplot(aes(x = class, y = conc, fill = class, color = class)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.2) +
  ggbeeswarm::geom_quasirandom(width = 0.25) + 
  xlab("") +
  ylab("Metabolite concentration") + 
  theme_minimal() +
  theme(legend.position = "none")
p
```

# Code {#code}

The following code was used to load, merge, and plot the (simulated) clinical trial data:

```{r show-code-1, echo=TRUE, eval=FALSE, ref.label='load-libraries'}
```

```{r show-code-2, echo=TRUE, eval=FALSE, ref.label='initial-data'}
```

```{r show-code-3, echo=TRUE, eval=FALSE, ref.label='new-data'}
```

```{r show-code-4, echo=TRUE, eval=FALSE, ref.label='combine-tables'}
```

```{r show-code-5, echo=TRUE, eval=FALSE, ref.label='plot-data'}
```

```{r show-table-1}
tmp <- cbind(df1[1:33,], rep('|', 33), 
             df1[34:66,], rep('|', 33), 
             df1[67:99,])
names(tmp) <- c('ID', 'Class', 'Conc', '|', 'ID', 'Class', 'Conc', 
                '|', 'ID', 'Class', 'Conc')

knitr::kable(tmp, booktabs = TRUE, 
             caption = "Initial subject data")
```

```{r show-table-2}
tmp <- cbind(df2[1:20,], rep('|', 20), 
             df2[21:40,], rep('|', 20), 
             df2[41:60,])
names(tmp) <- c('ID', 'Class', 'Conc', '|', 'ID', 'Class', 'Conc', 
                '|', 'ID', 'Class', 'Conc')

knitr::kable(tmp, booktabs = TRUE, 
             caption = "New subject data")
```

# Colophon

This manuscript was built at **`r format(Sys.time(), "%d %b %Y %H:%M:%S %Z")`** using the following computational environment and dependencies:

```{r colophon}
sessionInfo()
```

The current Git commit details are:

```{r}
# per Marwick, this line only executed if the user has installed {git2r} 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = '.'))
  git2r::commits(here::here())[[1]]
```

# References
